---
title: "skylight advanced"
author: "Koen Hufkens"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{skylight advanced}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  dpi = 300
)
```

## Cloud cover correction

The original implementation of the software allows for a correction of
illuminance based upon three cloud cover conditions. However, this is in all practical
sense a simple scaling of the data (see Janiczek and DeYoung 1987). You can either apply this scaling factor, a division by a number greater than 1 using the function itself, or apply it
post-hoc on the returned values. Below I show an example routine to use the internal scaling,
using ERA5 total cloud cover (tcc) data.

### Downloading cloud cover data

To provide our analysis with cloud cover data I rely on the BlueGreen Labs 
[`ecmwfr`](https://github.com/bluegreen-labs/ecmwfr) package. So, I first query some
total cloud cover data (0 - 1) covering most of Europe.

```{r eval = FALSE}
library(ecmwfr)

request <- list(
  product_type = "reanalysis",
  format = "netcdf",
  variable = "total_cloud_cover",
  year = "2021",
  month = "02",
  day = "01",
  time = c("00:00", "01:00", "02:00", "03:00", "04:00",
           "05:00", "06:00", "07:00", "08:00", "09:00",
           "10:00", "11:00", "12:00", "13:00", "14:00",
           "15:00", "16:00", "17:00", "18:00", "19:00",
           "20:00", "21:00", "22:00", "23:00"),
  area = c(70, -20, 33, 25),
  dataset_short_name = "reanalysis-era5-single-levels",
  target = "era5.nc"
)

wf_request(
  user = "xxxx",
  request = request,
  transfer = TRUE
)
```

Downloaded data can be read with the `terra` package. The above downloaded data is included as a demo
data set in the package. You can therefore use this included data in this workflow. Check out the `ecmwfr`
package to download your own custom data.

```{r echo = FALSE}
library(terra)

# read in data
era5 <- try(terra::rast(system.file(package = "skylight", "extdata/era5.nc")))

if(inherits(era5, "try-error")){
  process <- FALSE
} else {
  process <- TRUE
}
```


```{r eval = FALSE}
library(terra)

# read in data
era5 <- terra::rast(system.file(package = "skylight", "extdata/era5.nc"))

# use this command when downloading the data yourself
# era5 <- terra:rast(file.path(tempdir(),"era5.nc"))
```

However, the spatial format needs to be converted to a data frame for use
with the `skylight` package.

```{r eval = process}
library(dplyr)

# create a data frame with values
df <- era5 |>
  as.data.frame(xy = TRUE) |>
  rename(
    longitude = "x",
    latitude = "y"
  )

# add the original time stamp
df$date <- time(era5)

print(head(df))
```

```{r eval = process}
library(skylight)

# calculate sky illuminance values for
# a single date/time and location
df <- df |>
  mutate(
    # values of cloud cover lower than
    # 30% are considered clear conditions
    # for the skylight model - adjust tcc values
    tcc = ifelse(tcc <= 0.3, 0, 1),
    # rescale total cloud cover between 1 - 10
    # the acceptable range for skylight's
    # sky_condition parameter
    sky_condition = scales::rescale(df$tcc, to = c(1,10))
  ) |>
  # pipe into skylight
  skylight()

print(head(df))
```

I can now plot the sky illuminance as corrected for cloud cover using the internal scaling factor. Keep in mind that this is a rather ad-hoc solution. For proper results external empirical relationships should be established between the model response and measured illuminance values under varying cloud cover conditions.

```{r eval = process}
library(ggplot2)
library(rnaturalearth)

# country outlines
outlines <- rnaturalearth::ne_countries(returnclass = "sf")

ggplot(df) +
  geom_tile(
    aes(
      longitude,
      latitude,
      fill = log(total_illuminance)
    )
  ) +
  scale_fill_viridis_c(
    name = "Total illuminance (log(lux))"
  ) +
  geom_sf(
    data = outlines,
    colour = "white",
    fill = NA
    ) +
  coord_sf(
    xlim = c(-20, 25),
    ylim = c(33, 70)
  ) +
  labs(
    title = "Total illuminance with spatial variability\n due to cloud cover and location",
    y = "",
    x = ""
  ) +
  theme(
    legend.position = "bottom"
  )
```



